\section{Preliminaries}
\label{sec:prelim}

This section sets up the financial asset recommendation problem and the FAR-Trans benchmark that grounds our study, then the two-axis evaluation standard whose blind spot motivates the third axis this paper adds.

\subsection{The Financial Asset Recommendation Problem}
Financial asset recommendation (FAR) identifies and ranks financial securities such as stocks, bonds, and mutual funds for an investor according to their suitability. Suitability is two-sided. Investor-side signals include the customer's past transactions, declared risk tolerance, investment capacity, and personal goals, while market-side signals include asset returns, volatility, currency value, and inflation. A FAR system therefore draws on heterogeneous data sources: time-series pricing data, customer profile data, and historical investment transactions. Crucially, the investor-side half of suitability is a regulatory requirement under MiFID II, so a recommendation is only as good as its fit to the customer's risk profile, not its expected return or ranking quality alone.

\subsection{The FAR-Trans Benchmark}
Most prior FAR models were developed over proprietary or simulated data, which prevents fair cross-method comparison, and the only earlier public dataset lacked pricing data and asset identifiers, ruling out price-based approaches. FAR-Trans~\cite{sanzcruzado2024fartransinvestmentdatasetfinancial} closes this gap. It is the first public FAR dataset that pairs real asset pricing with real retail-investor transactions, collected from a large European financial institution and spanning January 2018 to November 2022. It provides daily closing prices, 228{,}913 Buy and 159{,}135 Sell transactions over 806 assets and 29{,}090 predominantly retail customers, and customer profile metadata including a declared MiFID II risk band per customer. Together these make both price-based and profile-based recommendation measurable on a single dataset.

Alongside the data, FAR-Trans establishes the field's evaluation standard by benchmarking 11 FAR algorithms. Two anchor our study because each leads on one axis of that benchmark: LightGCN~\cite{he2020lightgcnsimplifyingpoweringgraph}, a graph-based collaborative filter that propagates embeddings over the user-item bipartite graph and is the strongest ranking model, and Random Forest, a tree-based ensemble that ranks assets from tabular customer and asset features and is the strongest profitability model. The benchmark scores both over the top-10 ranked items on two axes: normalised Discounted Cumulative Gain (nDCG@10) for next-purchase ranking accuracy and Return on Investment (ROI@10) for realised profitability.

\subsection{The Suitability Blind Spot}
These two axes are not aligned on FAR-Trans. A subsequent study~\cite{10.1145/3780097} shows that transaction-based and profitability-based metrics are negatively correlated, so a model that wins on one tends to lose on the other, and concludes that ranking accuracy and realised return capture genuinely different objectives. Yet neither axis measures whether a recommendation aligns with the customer's declared MiFID II risk band. That regulatory suitability dimension is the blind spot this paper targets: we treat it as a third evaluation axis, risk suitability, and ask whether it can be balanced against return and accuracy rather than traded off. The next section makes risk suitability measurable and establishes that it is both prevalent enough to matter and economically consequential.
